On the Complexity of Best Arm Identification

### 2017-05-28 Complexity of Best Arm Identification/kl_complexity.Rmd
Add Track-and-Stop algorithm

This lower bound can be interpreted as a Minimax game, where the experimenter has to choose the share of samples he attributes to each arm in a way that is optimal given that nature will respond and choose the most difficult alternative model given his weights. Thus, to make the lower bound practical, it is a matter of finding the optimal weights $w^*(\bm{\mu})$.

>>>>

This lower bound can be interpreted as a Minimax game, where the experimenter has to choose the share of samples he attributes to each arm in a way that is optimal given that nature will respond and choose the most difficult alternative model given his weights. Thus, to make the lower bound practical, it is a matter of finding the optimal weights $w^*(\bm{\mu})$. This step did not occur in the previous lower bound, as the alternative models were made specific by changing the distribution of one arm with respect to the "border" between the two groups of arms.

In any case, the existence of the optimal weights suggests an algorithm that keeps track of the share of samples that is attributed to each arm and makes sure that it stays close to the optimum until some decision rule assures that the $\delta$-PAC condition is fulfilled.

And so by assuring the reader that there are ways to compute $w_a^*$ and $T^*$ given $\bm{\mu}$ (see chapter 2.2 in Garivier and Kaufmann (2016)), we can now introduce the *Track-and-Stop* strategy given in Garivier and Kaufmann (2016) that is able to match the lower bound by the intuition given above.

When going from theory to algorithm, the obvious problem is of course that we do no longer know $\bm{\mu}$ and thus cannot simply calculate the optimal weights. Instead, we are estimating the maximum likelihood estimates $\hat{\mu}(t) = (\hat{\mu}_1(t), \dots, \hat{\mu}_K(t))$ of the true mean at the current round $t$ and need to base our weights based on these. It is however problematic to rely on these sample means as we are inherently unsure about them. And so wrong estimates, which naturally occur most easily in the beginning when not many samples have been pulled, are prone to lead to wrong weights. Thus, if a wrong estimate $\hat{\mu}_a$ for arm $a$ leads to a small weight $w_a$, then neither the estimate nor the weight will be correct soon. Thus, Kaufmann and Garivier conclude that "a naive plug-in sampling rule sometimes fails". They propose to enforce a certain minimum level of exploration to make sure that arms that erroneously received a very small weight will be further explored regardless.

A very intuitive fix to this problem is the solution proposed in Lemma 8 of Garivier and Kaufmann (2016), with their $D-Tracking$ rule. Basically, whenever an arm has received a very low number of pulls $N_a(t)$ given the number of iterations so far, $t$ and the number of arms $K$ that have to share the samples, pull the arm that has received the smallest amount of samples so far. Else, pull the arm which number of pulls so far $N_a(t)$ is the furthest away from what its optimal share $t\cdot w^*_a(\hat{\bm{\mu}}(t))$ would be. Given this sampling rule, the authors can show in their Proposition 9 that the empirical proportion of draws converges in probability to the optimal weights given the true means.

When combining this sampling rule with an appropriate stopping rule, the authors are able to prove in Theorem 14 an upper bound on the expected sampling complexity $\mathbb{E}_{\mu}[\tau_{\delta}]$. The stopping rule ensures that the algorithm explores sufficiently to promise a risk smaller than $\delta$, while stopping as soon as possible. Given that the sampling rule is independent of this stopping rule, the rule boils down to a statistical test at every iteration and thus is a sequential test based on the current estimate of $\hat{\bm{\mu}}$. The Chernoff's Stopping Rule proposed in Garivier and Kaufmann (2016) calculates for each arm $a\in \mathcal{A}$ the Generalized Likelihood Ratio statistic with every arm $b \in \mathcal{A}^{\setminus a}$, which for $\hat{\mu}_a(t) \geq \hat{\mu}_b(t)$ can be written as

\begin{equation*}
Z_{a,b}(t) = - Z_{b,a} = N_a(t) \KL(\hat{\mu},\hat{\mu}_{a,b})) + N_b(t) \KL(\hat{\mu}_{b(t)}, \hat{\mu}_{a,b}),
\end{equation*}

\begin{equation*}
\hat{\mu}_{a,b}(t) := \frac{N_a(t)}{N_a(t) + N_b(t)}\hat{\mu}_a + \frac{N_b(t)}{N_a(t) + N_b(t)}\hat{\mu}_b.
\end{equation*}

Consequently, a sensible idea for a stopping rule would be that there has to be an arm $a \in \mathcal{A}$ for which $Z_{a,b} \geq 0 \forall b \in \mathcal{A}^{\setminus a}$ (which is always fulfilled if there is a uniquely best arm), and, additionally, we demand that for this arm the smallest of the $(K-1)$ test statistics has to be larger than an exploration threshold $\beta(t, \delta)$. Consequently, the final number of samples $\tau_{\delta}$ of an algorithm is given by:

\begin{equation*}
\tau_{\delta} = \inf \{t \in \mathbb{N}: \exists a \in \mathcal{A}, \forall b\in \mathcal{A}^{\setminus a}, Z_{a,b}(t) > \beta(t, \delta)\}
\end{equation*}

Demanding that the optimal exploration threshold increases in rounds $t$ (if we have more samples, require more certainty; also, punish for failing previous tests) and decreases in the risk $\delta$, and does so at a diminishing rate, the optimal threshold derived in Theorem 10 of Garivier and Kaufmann (2016) makes sense intuitively:

\begin{equation*}
\beta(t, \delta) = \log \Big(\frac{2t(K-1)}{\delta}\Big)
\end{equation*}.

### 2017-05-28 Complexity of Best Arm Identification/kl_complexity.tex
Add Track-and-Stop algorithm

This lower bound can be interpreted as a Minimax game, where the
experimenter has to choose the share of samples he attributes to each
arm in a way that is optimal given that nature will respond and choose
the most difficult alternative model given his weights. Thus, to make
the lower bound practical, it is a matter of finding the optimal weights
\(w^*(\bm{\mu})\).

>>>>

This lower bound can be interpreted as a Minimax game, where the
experimenter has to choose the share of samples he attributes to each
arm in a way that is optimal given that nature will respond and choose
the most difficult alternative model given his weights. Thus, to make
the lower bound practical, it is a matter of finding the optimal weights
\(w^*(\bm{\mu})\). This step did not occur in the previous lower bound,
as the alternative models were made specific by changing the
distribution of one arm with respect to the ``border'' between the two
groups of arms.

In any case, the existence of the optimal weights suggests an algorithm
that keeps track of the share of samples that is attributed to each arm
and makes sure that it stays close to the optimum until some decision
rule assures that the \(\delta\)-PAC condition is fulfilled.

And so by assuring the reader that there are ways to compute \(w_a^*\)
and \(T^*\) given \(\bm{\mu}\) (see chapter 2.2 in Garivier and Kaufmann
(2016)), we can now introduce the \emph{Track-and-Stop} strategy given
in Garivier and Kaufmann (2016) that is able to match the lower bound by
the intuition given above.

When going from theory to algorithm, the obvious problem is of course
that we do no longer know \(\bm{\mu}\) and thus cannot simply calculate
the optimal weights. Instead, we are estimating the maximum likelihood
estimates \(\hat{\mu}(t) = (\hat{\mu}_1(t), \dots, \hat{\mu}_K(t))\) of
the true mean at the current round \(t\) and need to base our weights
based on these. It is however problematic to rely on these sample means
as we are inherently unsure about them. And so wrong estimates, which
naturally occur most easily in the beginning when not many samples have
been pulled, are prone to lead to wrong weights. Thus, if a wrong
estimate \(\hat{\mu}_a\) for arm \(a\) leads to a small weight \(w_a\),
then neither the estimate nor the weight will be correct soon. Thus,
Kaufmann and Garivier conclude that ``a naive plug-in sampling rule
sometimes fails''. They propose to enforce a certain minimum level of
exploration to make sure that arms that erroneously received a very
small weight will be further explored regardless.

A very intuitive fix to this problem is the solution proposed in Lemma 8
of Garivier and Kaufmann (2016), with their \(D-Tracking\) rule.
Basically, whenever an arm has received a very low number of pulls
\(N_a(t)\) given the number of iterations so far, \(t\) and the number
of arms \(K\) that have to share the samples, pull the arm that has
received the smallest amount of samples so far. Else, pull the arm which
number of pulls so far \(N_a(t)\) is the furthest away from what its
optimal share \(t\cdot w^*_a(\hat{\bm{\mu}}(t))\) would be. Given this
sampling rule, the authors can show in their Proposition 9 that the
empirical proportion of draws converges in probability to the optimal
weights given the true means.

When combining this sampling rule with an appropriate stopping rule, the
authors are able to prove in Theorem 14 an upper bound on the expected
sampling complexity \(\mathbb{E}_{\mu}[\tau_{\delta}]\). The stopping
rule ensures that the algorithm explores sufficiently to promise a risk
smaller than \(\delta\), while stopping as soon as possible. Given that
the sampling rule is independent of this stopping rule, the rule boils
down to a statistical test at every iteration and thus is a sequential
test based on the current estimate of \(\hat{\bm{\mu}}\). The Chernoff's
Stopping Rule proposed in Garivier and Kaufmann (2016) calculates for
each arm \(a\in \mathcal{A}\) the Generalized Likelihood Ratio statistic
with every arm \(b \in \mathcal{A}^{\setminus a}\), which for
\(\hat{\mu}_a(t) \geq \hat{\mu}_b(t)\) can be written as

\begin{equation*}
Z_{a,b}(t) = - Z_{b,a} = N_a(t) \KL(\hat{\mu},\hat{\mu}_{a,b})) + N_b(t) \KL(\hat{\mu}_{b(t)}, \hat{\mu}_{a,b}),
\end{equation*}

\begin{equation*}
\hat{\mu}_{a,b}(t) := \frac{N_a(t)}{N_a(t) + N_b(t)}\hat{\mu}_a + \frac{N_b(t)}{N_a(t) + N_b(t)}\hat{\mu}_b.
\end{equation*}

Consequently, a sensible idea for a stopping rule would be that there
has to be an arm \(a \in \mathcal{A}\) for which
\(Z_{a,b} \geq 0 \forall b \in \mathcal{A}^{\setminus a}\) (which is
always fulfilled if there is a uniquely best arm), and, additionally, we
demand that for this arm the smallest of the \((K-1)\) test statistics
has to be larger than an exploration threshold \(\beta(t, \delta)\).
Consequently, the final number of samples \(\tau_{\delta}\) of an
algorithm is given by:

\begin{equation*}
\tau_{\delta} = \inf \{t \in \mathbb{N}: \exists a \in \mathcal{A}, \forall b\in \mathcal{A}^{\setminus a}, Z_{a,b}(t) > \beta(t, \delta)\}
\end{equation*}

Demanding that the optimal exploration threshold increases in rounds
\(t\) (if we have more samples, require more certainty; also, punish for
failing previous tests) and decreases in the risk \(\delta\), and does
so at a diminishing rate, the optimal threshold derived in Theorem 10 of
Garivier and Kaufmann (2016) makes sense intuitively:

\begin{equation*}
\beta(t, \delta) = \log \Big(\frac{2t(K-1)}{\delta}\Big)
\end{equation*}

.
